---
title: "ESM 244 Lab 6 Part 2: Rasters"
author: "Carlos Simms"
date: "2/21/2021"
output: html_document
---

### Attaching packages

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
library(tidyverse) # Default.
library(here) # To create file paths to data within different folders.
library(sf) # To read in spatial data and make simple features.
library(raster) # To work with raster data.
library(fasterize) # To make simple features and geometry.
```

### Reading in the raster data

NOTE: Working with raster data and images on the R Bren Server 3.6 will produce chunky, pixelated raster images on the .Rmd, but will produce clean, smooth raster images when knitted to an html.

```{r}
knitr::include_graphics(here("img", 
                             "landsat.png"))
```

### Loading in and inspecting the data

```{r}
landsat_file <- here("data", 
                     "Landsat7.tif")

ls_1 <- raster(landsat_file)
ls_1
```

```{r}
plot(ls_1)
```

```{r}
ls_2 <- raster(landsat_file, 
               band = 2)
ls_3 <- raster(landsat_file, 
               band = 3)
ls_4 <- raster(landsat_file, 
               band = 4)

ls_stack <- raster::stack(landsat_file) # To read all the layers in at once.
ls_stack
```

### Preparing the data by overwrighting the original layers

```{r}
ls_1 <- raster::aggregate(ls_1, 
                          fact = 3, 
                          fun = mean)
ls_2 <- raster::aggregate(ls_2, 
                          fact = 3, 
                          fun = mean)
ls_3 <- raster::aggregate(ls_3, 
                          fact = 3, 
                          fun = mean)
ls_4 <- raster::aggregate(ls_4, 
                          fact = 3, 
                          fun = mean)

plot(ls_1, 
     col = hcl.colors(n = 100, 
                      palette = 'Blues 2')) # To base a color scheme on a single palette.
plot(ls_2, 
     col = hcl.colors(n = 100, 
                      palette = 'Greens 2'))
plot(ls_3, 
     col = hcl.colors(n = 100, 
                      palette = 'Reds 2'))
plot(ls_4, 
     col = hcl.colors(n = 100, 
                      palette = 'Reds 2'))
```

```{r}
sbc_rast <- raster(here("data", 
                        "county.tif"))

plot(sbc_rast)

plot(ls_3)

# ERROR Code in Lab Playthrough:

# mask(ls_3, sbc_rast) %>% plot()

# ls_3 <- mask(ls_3, sbc_rast)

# ls_4 <- mask(ls_4, sbc_rast)
```

### Working with raster algebra

```{r}
vec1 <- 1:5
vec1
vec1*2
vec1^2

ls_3
ls_3*2
log(ls_3)
plot(ls_3); plot(log(ls_3)) # To compare.
```

```{r}
vec2 <- 6:10
vec1+vec2
ls_3+ls_4
```

### Using 'raster::calc()'

```{r}
ls_stack <- stack(ls_1, 
                  ls_2, 
                  ls_3, 
                  ls_4)

ls_mean <- raster::calc(ls_stack, 
                        fun = mean, 
                        na.rm = FALSE)

plot(ls_mean)
```

### Analysis: NDVI

```{r}
knitr::include_graphics(here("img", 
                             "spectrum.png"))

knitr::include_graphics(here("img", 
                             "ir_photo.jpg"))
```

### NDVI equation

$$NDVI = \frac{NIR - Red}{NIR + Red}$$

```{r}
ndvi <- (ls_4 - ls_3) / (ls_4 + ls_3)

plot(ndvi, 
     col = hcl.colors(100, 
                      palette = 'Grays'))

is_forest <- function(x, 
                      thresh = 0.3) {
  y <- ifelse(x >= thresh, 
              1, 
              NA)
  return(y)
}

forest <- calc(ndvi, 
               fun = is_forest)

plot(forest, 
     col = 'green')
```

### 'ggplot' and rasters

```{r}
ndvi_df <- raster::rasterToPoints(ndvi) %>% 
  as.data.frame()

forest_df <- raster::rasterToPoints(forest) %>% 
  as.data.frame()

ggplot(data = ndvi_df, 
       aes(x = x, 
           y = y, 
           fill = layer)) + 
  geom_raster() + 
  geom_raster(data = forest_df, 
              fill = "green") + 
  coord_sf(expand = 0) + 
  scale_fill_gradient(low = "black", 
                      high = "white") + 
  theme_void() + 
  theme(panel.background = element_rect(fill = "slateblue4"))
```
